### 2025-03-12 index.qmd
added placeholders for analyses

# README

Gender in authors in music psychology

What is the gender distribution is music psychology specialist journals?

The public-facing version of the repository is available at
[https://tuomaseerola.github.io/gender_in_music_psych/](https://tuomaseerola.github.io/gender_in_music_psych/)

## Data

Data comes from this source and ....

## Preprocessing

To attribute the gender for the authors, we used gender API [analysis/gender_attribution.qmd](analysis/gender_attribution.qmd).

## Analysis

Data analysis is covered in [analysis/analysis.qmd](analysis/analysis.qmd) document.

## Manuscript

The study report is available [manuscript/manuscript.qmd](manuscript/manuscript.qmd) document.