### 2025-08-05 analysis.qmd
Updated the analysis and the manuscript and the data

# Analyse gender in music psychology

T. Eerola, 1/8/2025

Gender analysis of the music psychology journals since 2000 (Musicae Scientiae, Psychology of Music, Music Perception, Journal of New Music Research, and Music & Science).

## Required libraries

```{r}
#| id: libraries
#| warning: false
#| eval: true
#| echo: false
#| message: false
library(tidyverse)
library(stringr)
library(ggplot2)
library(bib2df)
library(countrycode)
library(modelsummary)
#library(DescTools)
library(easystats)
library(here)
```

## Load and preprocess data

```{r}
#| id: preprocess
#| eval: true
#| output: asis
source('scripts/load_scopus_datasets.R') # Creates full_names dataset
source('scripts/attribute_country.R') # Add country affiliations
source('scripts/attribute_gender.R') # Adds gender attributions (precalculated with API)
source('scripts/create_keys.R') # Add keys for authors and study+author
source('scripts/clean_citations_and_OA.R') # Process citations and OA status
source('scripts/attribute_gender_diagnostics.R') # Process citations and OA status
source('scripts/export_gender_data.R')
```

## Summarise

```{r}
#| id: summarise
#| eval: true
#| output: asis
#| warning: false
#| messages: false
source('scripts/summarise_gender.R') # different summaries
```

## Quantify

```{r}
#| id: quantify
#| eval: true
#| warning: false
#| messages: false
#| output: asis
source('scripts/quantify_authorship.R') # use Odds
source('scripts/citations.R')           # Citations and gender
print(knitr::kable(citestats_all, digits = 2, caption = 'Citations across all authors'))
print(knitr::kable(stats_all))

source('scripts/open_access.R')         # OA and gender
print(knitr::kable(author_OA, digits = 2, caption = 'Open access across all authors'))
```

## Geographical

```{r}
#| id: geographical
#| eval: true
#| output: asis

source('scripts/country_table.R') # OK
source('scripts/visualise_gender.R') # OK

```

```{r}
#| id: geographical2
#| eval: true
#| echo: true
#| output: asis
#source('scripts/table_continent.R') # IN progress
```

Table: Female author odds ratios across continents.

|Continent |    N| Odds_ratio| CI_low| CI_high|
|:---------|----:|----------:|------:|-------:|
|Europe    | 4985|       0.82|   0.75|    0.89|
|Americas  | 2356|       1.12|   1.02|    1.23|
|Asia      |  931|       0.89|   0.78|    1.02|
|Oceania   |  742|       1.66|   1.43|    1.93|
|Africa    |   52|       0.91|   0.52|    1.59|