### 2025-08-05 index.qmd

# Gender in music psychology

What is the gender distribution of publishing authors is music psychology? Has there been a change over the years in authorship positions (first, coauthor, last) and how is this discipline compared to for instance to psychology?

The public-facing version of the repository is available at
[https://tuomaseerola.github.io/gender_in_music_psych/](https://tuomaseerola.github.io/gender_in_music_psych/)

## Data

Data comes from this Scopus with the specific syntax: `SRCTITLE ( JOURNAL NAME ) AND PUBYEAR > 1999`. We focussed on 5 highest ranked journals in Music Psychology, namely _Musicae Scientiae_, _Psychology of Music_, _Music Perception_, _Journal of New Music Research_, and _Music & Science_. To attribute the gender for the authors, we used genderize API from the extracted first names. We manually checked a subset of the data and attributed some of those names that the API could not resolve.

## Analysis

Data analysis is covered in [analysis.qmd](analysis.qmd) document.

## Manuscript

The study report is available [manuscript/manuscript.qmd](manuscript/manuscript.qmd) document.