### 2025-08-05 index.qmd
Updated the analysis and the manuscript and the data

# Gender in music psychology

What is the gender distribution of publishing authors is music psychology? Has there been a change over the years in authorship positions (first, coauthor, last) and how is this discipline compared to for instance to psychology?

The public-facing version of the repository is available at
[https://tuomaseerola.github.io/gender_in_music_psych/](https://tuomaseerola.github.io/gender_in_music_psych/)

## Data

Data comes from this Scopus with the specific syntax: `SRCTITLE ( JOURNAL NAME ) AND PUBYEAR > 1999`. We focussed on 5 journals in Music Psychology, namely _Musicae Scientiae_, _Psychology of Music_, _Music Perception_, _Journal of New Music Research_, and _Music & Science_. To attribute the gender for the authors, we used [genderize API](https://genderize.io/). The details of the process are given in the manuscript.

## Analysis

Data analysis steps are covered in the [analysis.qmd](analysis.qmd) document.

## Manuscript

The study report is available at [manuscript/manuscript.qmd](manuscript/manuscript.qmd) document.