### 2025-08-26 manuscript/_figure2.qmd

```{r}
#| label: fig2
#| fig-width: 9
#| fig-height: 4
#| echo: false
#| warning: false
#| message: false
#| fig.cap: > 
#|   Authorship types across author counts.

source(here("scripts","load_gender_data.R"))
library(ggplot2,quietly = TRUE)
library(RColorBrewer)

df$author_arbitrary <- cut(
  df$author_order,
  include.lowest = TRUE,
  breaks = c(0, 1, 2, 3, 4, 5, 6, 7, 8,40),
  labels = c("1", "2", "3", "4", "5", "6", "7", "8","9+")
)

df <- dplyr::mutate(dplyr::group_by(df, BIBTEXKEY), author_N = max(author_order))
#table(df$author_N)

df$author_N_cropped <- cut(
  df$author_N,
  include.lowest = TRUE,
  breaks = c(0, 1, 2, 3, 4, 5, 6, 40),
  labels = c("1", "2", "3", "4", "5", "6","7+")
)
#table(df$author_N_cropped)

#### classify order: First author -------
df$authortype <- 'Other'
df$authortype[df$author_order == 1] <- 'First'
df$authortype[df$author_last == 1] <- 'Other'
df$authortype[df$author_single == 1] <- 'Other'
#table(df$authortype,df$Gender)

output <- NULL
U<-levels(df$author_N_cropped)
for (k in 2:length(U)){
  tmp<-dplyr::filter(df,author_N_cropped==U[k])
#  t <- table(tmp$authortype,tmp$Gender)
  t <- table(
    factor(tmp$authortype, levels = c("First", "Other")),
    factor(tmp$Gender,levels = c("female","male"))
  )
  output <- rbind(output,effectsize::oddsratio(t))
}

output$author_N_cropped <- U[2:length(U)]
first <- data.frame(output)
#first

#### classify order: Coauthor ----------
df$authortype <- 'Coauthor'
df$authortype[df$author_order == 1] <- 'Other'
df$authortype[df$author_last == 1] <- 'Other'
df$authortype[df$author_single == 1] <- 'Other'
#table(df$authortype,df$Gender)

output <- NULL
U<-levels(df$author_N_cropped)
for (k in 3:length(U)){
  tmp<-dplyr::filter(df,author_N_cropped==U[k])
#  t <- table(tmp$authortype,tmp$Gender)
  t <- table(
    factor(tmp$authortype, levels = c("Coauthor", "Other")),
    factor(tmp$Gender,levels = c("female","male"))
  )

  output <- rbind(output,effectsize::oddsratio(t))
}

output$author_N_cropped <- U[3:length(U)]
coauthor <- data.frame(output)

#### classify order: Last ----------
df$authortype <- 'Other'
df$authortype[df$author_order == 1] <- 'Other'
df$authortype[df$author_last == 1] <- 'Last'
df$authortype[df$author_single == 1] <- 'Other'

output <- NULL
U<-levels(df$author_N_cropped)
for (k in 2:length(U)){
  tmp<-dplyr::filter(df,author_N_cropped==U[k])
  #t <- table(tmp$authortype,tmp$Gender)
  t <- table(
    factor(tmp$authortype, levels = c("Last", "Other")),
    factor(tmp$Gender,levels = c("female","male"))
  )

  output <- rbind(output,effectsize::oddsratio(t))
}

output$author_N_cropped <- U[2:length(U)]
last <- data.frame(output)

first$Type<-'First'
coauthor$Type<-'Co'
last$Type<-'Last'

author_N<-rbind(first,coauthor,last)
author_N$Type<-factor(author_N$Type,levels=c("First","Co","Last"))

# Can we count this for single-authored papers?
tmp<-dplyr::filter(df,author_N==1)
t<-table(tmp$Gender)
tmp2<-dplyr::filter(df,author_N>1)
t2<-table(tmp2$Gender)
T<-data.frame(rbind(single=t,multi=t2))
single<-data.frame(effectsize::oddsratio(T),author_N_cropped=1,Type="Single")
author_N <- rbind(single,author_N)
author_N$Type<-factor(author_N$Type,levels=c("Single","First","Co","Last"))

custom_palette <- brewer.pal(n = 4, name = "Set1")
custom_palette<-c(custom_palette[4],custom_palette[1:3])
fig2 <- ggplot(author_N,aes(x=author_N_cropped,y=Odds_ratio,color=Type))+
  geom_hline(yintercept=1, linetype="dashed", color = "black") +
  geom_point(position = position_dodge(width = 0.4),shape=15,size=3)+
  geom_errorbar(aes(ymin=CI_low,ymax=CI_high),width=0.2,position = position_dodge(width = 0.4))+
  scale_color_manual(name="Authorship\ntype",values = custom_palette) +
  xlab('Author N')+
  ylab('Female Authorship Odds Ratio')+
  theme_classic(base_size = 15)
print(fig2)
```